### 2020-12-11 doc/manuscript/Dow_manuscript_Nature.Rmd
improved abstract

(*word limit 200; this is currently 266*)

As the climate changes, warmer spring temperatures are causing earlier leaf-out in temperate deciduous forests, along with earlier initiation of the annual period where these ecosystems are net sinks of carbon dioxide (CO~2~).
As a result, there is a tendency towards increased growing season length and annual CO~2~ sequestration.
However, little is known about how spring temperatures affect tree stem growth, which consumes a modest portion of the carbon fixed annually but is critical in that wood has a long residence time in the ecosystem.
We need to know how the timing and rate of woody growth are affected by warming spring temperatures.
Here, we show that in two temperate deciduous forests warmer maximum temperatures in the spring shift growth milestones earlier in the year, but have only a modest effect on peak growing season length and little effect on annual growth.
This contrasts with the expectation *in most models* that an earlier start to growing season would significantly increase the rate of biomass accumulation and thereby enhance the CO~2~ sink of temperate deciduous forests.
It is, however, consistent with most tree-ring studies in temperate deciduous forests showing that annual growth tends to be more sensitive to climatic conditions during the peak growing season than in the spring[@helcoski_growing_2019], with eddy-covariance records showing that annual CO~2~ sequestration is more strongly linked to conditions in the peak growing season than to growing season length[@fu_maximum_2019], and with evidence that forest CO~2~ sequestration is sink-limited[@zani_increased_2020].
Thus, our findings indicate that warming spring conditions are unlikely to impact temperate deciduous forest growth sufficiently to produce any substantial negative climate change feedback.

@@

(*most or all of this should be cut-- first paragraph too detailed for Nature, and we want minimal comparison of ring vs diffuse porous*)
Growth milestones were achieved earlier at SCBI than at Harvard Forest (average difference: #-# days) and differed between ring- and diffuse-porous species (Fig. `r growth_curves_all`, Table `r mean_growth_parameters_table`).
Across all trees and years, $DOY_{25}$ occurred on day 139 (May 20) at SCBI and day 151 (June 1) at Harvard Forest, while $DOY_{75}$ occurred on day 139 (May 20) at SCBI and day 151 (June 1) at Harvard Forest.
Ring-porous species began growing earlier, reaching the $DOY_{25}$ benchmark earlier (by `r 154-123` days at SCBI and `r 164-132` at Harvard Forest), and their growth was spread out over a longer growing season (average $L_{PGS}$ 21 and 19 days longer at SCBI and Harvard Forest, respectively). 
Growing season length was similar across sites, with $L_{PGS}$ being, on average, only two days longer at SCBI for ring-porous species and <1 day longer for diffuse-porous species.

>>>>

(*word limit 200. italic=tentative*)

As the climate changes, warmer spring temperatures are causing earlier leaf-out[REFS] and commencement of net carbon dioxide (CO~2~) sequestration[REFS] in temperate deciduous forests, resulting in a tendency towards increased growing season length[REFS] and annual CO~2~ sequestration[REFS].
However, little is known about how spring temperatures affect tree stem growth, which consumes a modest portion of the carbon fixed annually[REFS] but is the longest-lived carbon sink[REFS].
Here, we show that in two temperate deciduous forests, warmer maximum spring temperatures consistently shifted *both leaf phenology* and woody growth milestones earlier but had mixed, weak effects on peak growing season length, maximum daily growth rates, and annual growth.
*The latter finding is confirmed on the centennial scale by tree-ring records across # forests.*
These findings are consistent with *~~most tree-ring studies in temperate deciduous forests showing that annual growth tends to be more sensitive to climatic conditions during the peak growing season than in the spring[REFS], with~~* eddy-covariance records showing that annual CO~2~ sequestration is more strongly linked to conditions in the peak growing season than to growing season length[@fu_maximum_2019], with observations that warm springs result in earlier leaf senescence[@zohner_ongoing_2019,@zani_increased_2020], and with evidence that forest CO~2~ sequestration is sink-limited[@zani_increased_2020].
They contrast, however, with the expectation *in most models[@zani_increased_2020]* that an earlier start to growing season would significantly increase the rate of biomass accumulation and thereby enhance the CO~2~ sink of temperate deciduous forests, implying that warming spring conditions are unlikely to impact temperate deciduous forest growth sufficiently to produce any substantial negative climate change feedback.

@@

*(we will need to add leaf phenology to this paragraph)*

**Growth milestones were achieved earlier at SCBI than at Harvard Forest (average difference: #-# days) and differed between ring- and diffuse-porous species (Fig. `r growth_curves_all`, Table `r mean_growth_parameters_table`).**
Across all trees and years, $DOY_{25}$ occurred on day 139 (May 20) at SCBI and day 151 (June 1) at Harvard Forest, while $DOY_{75}$ occurred on day 185 (July 5) at SCBI and day 194 (July 14) at Harvard Forest.
Ring-porous species began growing earlier, reaching the $DOY_{25}$ benchmark earlier (by `r 154-123` days at SCBI and `r 164-132` at Harvard Forest), and their growth was spread out over a longer growing season (average $L_{PGS}$ 21 and 19 days longer at SCBI and Harvard Forest, respectively). 
Growing season length was similar across sites, with $L_{PGS}$ being, on average, only two days longer at SCBI for ring-porous species and <1 day longer for diffuse-porous species.
